Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance Settings

### 2019-01-29 Manuscript/manuscript.Rmd
editted methods

We first assessed predictor type performance using random forest classification algorithms by utilizing each set of predictor types as the feature space, in addition to a model with all combined types. Models were evaluated across each combination of resampling techniques applied to the training set, for each cell line. Performance was based on the AUPRC

To assess the performance for each predictor type, we performed individual random forest classification algorithms utilizing each set of predictor types as the feature space, in addition to a model with all combined types,. Precision-Recall (PR) Curves, and subsequent area under the PR curves (AUPRC), were used to evaluate which predictor type contributed to the best model performance. Recursive feature elimination (RFE) was then used to determine the optimal number of features for models within each set of predictor types. In doing so, random forest models were used to assign weights to a set features of a given size. The optimal number of features was determined by the AUPRC of the corresponding model in order to reduce the feature space accordingly.

For the task of identifying the functional genomic elements that were most predictive of the formation of TAD boundaries, we evaluated the variable importances associated with each predictor of the random forest model. For each tree, the prediction accuracy on the out-of-bag portion of the data was recorded. Then the same was done after permuting each predictor variable. The difference between the two accuracies was then averaged over all trees, and normalized by the standard error [@kuhn2012caret].

>>>>

We first assessed predictor type performance using random forest classification algorithms by utilizing each set of predictor types as the feature space, in addition to a model with all combined types. Models were evaluated across each combination of resampling technique applied to the training set, for each cell line. Performance was based on the AUPRC. Recursive feature elimination (RFE) was then used to determine the optimal number of features for models with the best performing predictor type, across resampling techniques and cell lines. In doing so, random forest models were again used to recursively assign weights to a set of features of a given size. For computational efficiency we evaluated feature subset sizes in descending order of powers of 2. The optimal number of features was determined by the AUPRC of the corresponding model. Optimal performance was assessed and the feature spaced was reduced accordingly.

For the task of identifying the functional genomic elements that were most predictive of the formation of TAD boundaries, we evaluated the variable importances associated with each predictor of the random forest models associated with each resampling technique. For each model and within each tree, the prediction accuracy on the out-of-bag portion of the data was recorded. Then the same was done after permuting each predictor variable. The difference between the two accuracies was then averaged over all trees, and normalized by the standard error. All random forest algorithms were implemented in R using the `caret` package version 6.0 [@kuhn2012caret].
